---
title: "STATS 101 Project"
author: "Sherry Shen, Lisa Zhong, Selina Zhang"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
set.seed(123)     #setting seed
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
print("trial")
```

## Splitting Data
```{r}
data <- read.csv("ames2000_NAfix.csv")
smp_size = floor(0.5 * nrow(data))

train_Index = sample(seq_len(nrow(data)), size = smp_size)
trainData <- data[train_Index,]
testData <- data[-train_Index,]

# data$MS.SubClass = factor(data$MS.SubClass)
# data$MS.Zoning = factor(data$MS.Zoning)
# data$Street = factor(data$Street)
# data$Alley = factor(data$Alley)
# data$Lot.Shape = factor(data$Lot.Shape, ordered = T) # ordinal
# data$Land.Contour = factor(data$Land.Contour) 
# data$Utilities = factor(data$Utilities, ordered = T)
# data$Lot.Config = factor(data$Lot.Config)
# data$Land.Slope = factor(data$Land.Slope, ordered = T)
# data$Neighborhood = factor(data$Neighborhood)
# data$Condition.1 = factor(data$Condition.1)
# data$Condition.2 = factor(data$Condition.2)
# data$Bldg.Type = factor(data$Bldg.Type)
# data$House.Style = factor(data$House.Style)
# data$Overall.Qual = factor(data$Overall.Qual, ordered = T)
# data$Overall.Cond = factor(data$Overall.Cond, ordered = T)
# data$Roof.Style = factor(data$Roof.Style)
# data$Roof.Matl = factor(data$Roof.Matl)
# data$Exterior.1st = factor(data$Exterior.1st)
# data$Exterior.2nd = factor(data$Exterior.2nd)
# data$Mas.Vnr.Type = factor(data$Mas.Vnr.Type)
# data$Mas.Vnr.Area = as.numeric(data$Mas.Vnr.Area)
# data$Exter.Qual = factor(data$Exter.Qual, ordered = T)
# data$Exter.Cond = factor(data$Exter.Cond, ordered = T)
# data$Foundation = factor(data$Foundation)
# data$Bsmt.Qual = factor(data$Bsmt.Qual, ordered = T)
# data$Bsmt.Cond = factor(data$Bsmt.Cond, ordered = T)
```
## The Study
```{r}
nrow(data)
```

What field does the data come from?
The data came from American Statistical Association, specifically from Ames, Iowa Assessor’s Office. It describes different features for "individual residential properties sold in Ames, IA from 2006 to 2010" (Stanford).

What are the goals of the study? Are there any effects of particular interest?
Our population of interest is the individual real estates sold in Ames. The Population Parameter is the sale price of the real estates. 

How might these goals be answered, i.e. tests / confidence intervals?
