---
title: "STATS 101 Project"
author: "Sherry Shen, Lisa Zhong, Selina Zhang"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
set.seed(123)     #setting seed
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
print("trial")
```

##Conducting PCA
```{r}
data <- read.csv("ames2000_NAfix.csv", stringsAsFactors = TRUE)

unique(data$Utilities) 
data <- data[,-9]

library("factoextra")

dataM = as.matrix(data)
dataMat = as.matrix(data) %*% t(dataM)
## standardizing the data
dataMat = scale(dataMat)
mean(dataMat[, 1])
sd(dataMat[, 1])
pca_data = princomp(dataMat)
#summary(pca_x)
pca_data$loadings[, 1:2]
fviz_eig(pca_data, addlabels = TRUE)
plot(pca_data$loadings[, 1],
     pca_data$loadings[, 2], 
     xlab='loadings 1', 
     ylab='loadings 2',
     main='PCA on data')


```


## Splitting Data

```{r}

smp_size = floor(0.5 * nrow(data))

train_Index = sample(seq_len(nrow(data)), size = smp_size)
trainData <- data[train_Index, ] # removing Utilities
testData <- data[-train_Index, ] # removing Utilities

# data$Lot.Shape = factor(data$Lot.Shape, ordered = T) # ordinal
# data$Utilities = factor(data$Utilities, ordered = T)
# data$Land.Slope = factor(data$Land.Slope, ordered = T)
# data$Overall.Qual = factor(data$Overall.Qual, ordered = T)
# data$Overall.Cond = factor(data$Overall.Cond, ordered = T)
# data$Exter.Qual = factor(data$Exter.Qual, ordered = T)
# data$Exter.Cond = factor(data$Exter.Cond, ordered = T)
# data$Bsmt.Qual = factor(data$Bsmt.Qual, ordered = T)
# data$Bsmt.Cond = factor(data$Bsmt.Cond, ordered = T)
```

## The Study

```{r}
head(data)
dim(data)
summary(data)

```

**What field does the data come from?** The data came from American Statistical Association, specifically from Ames, Iowa Assessor's Office. It describes different features for "individual residential properties sold in Ames, IA from 2006 to 2010" (Stanford).

**What are the goals of the study? Are there any effects of particular interest?** Our population of interest is the individual real estates sold in Ames. The Population Parameter is the sale price of the real estates. 

**How might these goals be answered, i.e. tests / confidence intervals?**

## The Data
```{r}

```